### 2012-06-20 intention motor representation with corrado sinigaglia/intention and motor action.ppr.corrected.tex
proof reading nearly done

How does representing an outcome differ from representing merely kinematic or dynamic features of action? How, for instance, does a motor representation of grasping (assuming for now that such a thing exists) differ from a representation of a sequence of joint displacements? First, a motor representation of grasping captures something common to many different sequences of joint displacements and postures involving a variety of effectors. To illustrate, an agent might grasp an object with her hand, with her mouth, with normal pliers (where grasping requires closing the hand), or with reverse pliers (where grasping requires opening the hand). A motor representation of grasping identifies something common to all these cases. Second, and conversely, a motor representation of grasping potentially distinguishes between the same sequence of joint displacements in different contexts. For instance, how a grasping action is represented may depend on a relatively distal outcome: on whether (say) it results in the object grasped being eaten or placed. Also the joint displacements which realise grasping in one context might in another context realise a different action, such as scratching; or they might fail to realise any action at all (because there is no target object, say). Third, a motor representation of an action outcome locates the action within a hierarchical structure. To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.\footnote{On the notion of action outcome in motor representation, see \citet{gallese:2000_inner}, \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}

Why accept that there are motor representations of action outcomes? The first step is to consider evidence that motor processes carry information about action outcomes. For any given marker of motor processing (such as a pattern of neuronal discharge or motor-evoked potentials), how can we test whether that marker carries information about action outcomes? The basic principle is straightforward: vary kinematic and dynamic features while holding constant an action outcome; and, conversely, vary action outcomes while holding kinematic and dynamic features constant. In practice researchers have devised many ingenious ways to achieve this. In order to vary kinematic and dynamic features while holding action outcomes constant, in some studies a single action outcome is achieved using different effectors, hand, mouth or foot, say \citep{rizzolatti:1988_functional,Rizzolatti:2001ug,cattaneo:2010_state-dependent}. A variation on this approach is to contrast performing a grasping action with different tools, so that the same action outcome might require closing or opening the hand depending on the tool used \citep{umilta:2008pliers,cattaneo:2009_representation,rochat:2010_responses}. In order to vary action outcome while holding kinematic and dynamic features constant, researchers have contrasted grasping movements with different distal outcomes such as eating and placing \citep{Fogassi:2005nf,bonini:2010_ventral,cattaneo:2007_impairment}. Another approach is to contrast the same grasping movements performed in the presence or manifest absence of a target object \citep{Umilta:2001zr,villiger:2010_activity}. A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}. In each of these cases there is evidence that some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.\footnote{ Of course some researchers have raised doubts concerning some of the evidence that there are markers of motor processing which carry information about action outcomes (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).  On balance, however, the evidence supports this conclusion.}

@@

So far we have argued that motor processes involve representations of action outcomes. It remains to show that such representations ground purposive actions. But this is hardly a further step. How do intentions ground the purposiveness of actions?  On any standard view, an intention represents an outcome, causes an action, and does so in a way that would normally facilitate the outcome's occurrence. Similarly, motor representations of outcomes represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented. To say that motor representations do all this is one way of making precise the metaphor involved in saying that purposive actions are directed to outcomes.  Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness (as \citealp{pacherie:2008_action}, pp.\ 189-90 has also argued).  So given the functional roles and representational objects of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.

@@

We have been comparing actually seeing with imagining seeing and actually acting with imagining acting as a first step towards arguing that visual and motor representations differ in format (which in turn will be background for an argument that motor representation and intention differ in format).  This claim could not easily be established by comparing actually seeing with actually acting because performance differences between these might be explained by bodily or environmental factors only distantly related to the representations involved. By contrast, comparing performance in imagining seeing and imagining acting does provide reasons to conclude that visual and motor representations differ in format. To see why, contrast imagining rotating a ball with imagining seeing a ball rotating.%
\footnote{Imagining acting without also imagining seeing may be difficult in practice, and conversely; it may also sometimes be difficult to distinguish imagining acting from imagining seeing \citep[as][p.\ 170 suggest]{currie:1997_mental}. However ordinary subjects can separate the two well enough to confirm predictions about their differences \citep[see, e.g.,][]{kosslyn:2001_imagining}.
}
As already mentioned, how quickly the former can be done is a function of how long it would take the agent to rotate the ball, whereas how quickly the latter can be done depends on how rapidly the ball can rotate and still be perceived as rotating. Similarly, we mentioned that factors making actually acting more effortful also make imagining acting more effortful. For instance, in some cases rotating a ball clockwise is easier than rotating it anti-clockwise, and so is imagining a ball rotate.  By contrast, the effort involved in actually seeing or imagining seeing a ball rotate does not similarly differ depending on direction.  These and other performance differences are plausibly a consequence of a difference in format between motor and visual representations.

@@

The idea that some demonstrative concepts refer to actions by deferring to motor representations immediately raises a question. Could a demonstrative concept really defer to a motor representation? It seems clear that we can't select a motor representation to defer to in the same way that we can select a map when we say `Follow this route!'. After all, motor representations are not things we can point to with our hands. Nevertheless, it does seem that motor representations are available in some sense. To start with an analogy, consider pantomiming an action to yourself. You are rehearsing part of an operation which involves precisely grasping a delicate structure with some tweezers. Just as someone might point to a map and say `Follow that route!', so also they could point to your pantomime and say `Do that!'. In our analogy, the pantomime stands in for a motor representation of the action.  The demonstrative in `Do that!' refers to an action by deferring to the pantomime. Of course this analogy doesn't show that demonstrative concepts can defer to motor representations (at least not unless pantomimes are motor representations of actions). But now consider purely mental pantomime---that is, phenomenologically action-like imagination. One might use this kind of imagination to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!' It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in imagining acting.

But is this really possible? Someone might object that in appealing to imagining acting we are sneaking intentions in through the back door. How can we be sure that it is ever really a motor representation rather than an intention that one defers to in thinking `Do that!'?  Contrast two cases of phenomenologically action-like imagination, both involving a tool. In the first case, imagine grasping an object with the tool; and in the second case, imagine releasing the same object with the tool. The two cases are different but in what does their difference consist? Could it be a difference in intention that explains how they differ?  Since it is possible to grasp without intending to grasp, it is surely also possible to imagine grasping without imagining intending to grasping.  And note that to imagine grasping without imagining intending to grasping does not necessarily require one to imagine non-intentionally grasping. (It is easy to miss this point by confusing it with the issue of whether the imagining must itself be intentional. Whether or not it is possible to imagine grasping or releasing without intending to imagine so acting, one can certainly imagine grasping or releasing without  imagining intending to so act.) So to rule out the possibility that the two imaginings differ with respect to intention, let both be neutral with respect to intention. Is the difference between the two cases then due to differing movements or patterns of muscle contraction? To rule this possibility out, let the movements and muscle contractions involved in both cases be as similar as possible: let the difference between grasping and releasing the object be a matter of how the tool is configured rather than of how your body moves. With the contrast cases elaborated by these two stipulations, it is plausible that the difference between the two imaginings will be due to the different motor representations involved in grasping and releasing. So in these cases, the thought `Do that!' will refer to grasping (or releasing) by deferring to a motor representation of grasping (or of releasing). It follows that demonstrative concepts can refer to actions by deferring to motor representations.

@@

At this point we should acknowledge a complication.  Above we argued that experience is needed for the acquisition of demonstrative concepts of ways of acting but not for their continued use in practical reasoning.  This matters because forming an intention does not always involve newly experiencing a way of acting. However, things are more complicated concerning what to act on. Where an action involves particular things to be acted on, agents will not generally have past experiences of these.  Accordingly, demonstrative reference by deference to motor representation requires that some motor representations of particular targets for action are associated with experiences of those targets.  When we spontaneously intend to act on an object of which we have no previous experience (and perhaps in other cases too), our experience of that object would have to depend in part on motor representations of it.  There is indeed some evidence for the ubiquity of motor aspects of experience (and also evidence that motor representation does not always modulate how objects are experienced, but that is not relevant here).  For example, temporarily changing subjects' motor abilities by artificially extending their reach systematically affects their judgements of how far away objects are \citep{linkenauger:2009_effects, costantini2011tool}. Given that such judgements are based on experience,\footnote{
It may be objected that these judgements could reflect non-experiential expectations; \citet[pp.\ 203-4]{witt:2011_action} reviews evidence against this possibility.
} 
this illustrates one way in which experience may be shaped in part by motor representation of objects.  There is also evidence that experiential judgements of size can be influenced by ability \citep[e.g.][]{witt2009kicking}.  It is plausible, then, that motor representation of objects modulates experiences of them.  This indicates that it is possible to refer not only to ways of acting by also to what to act on by deferring to motor representations.

>>>>

How does representing an outcome differ from representing merely kinematic or dynamic features of action? 
How, for instance, does a motor representation of grasping (assuming for now that such a thing exists) differ from a representation of a sequence of joint displacements? 
First, a motor representation of grasping captures something common to many different sequences of joint displacements and postures involving a variety of effectors. 
To illustrate, an agent might grasp an object with her hand, with her mouth, with normal pliers (where grasping requires closing the hand), or with reverse pliers (where grasping requires opening the hand). 
A motor representation of grasping identifies something common to all these cases. 
Second, and conversely, a motor representation of grasping potentially distinguishes between the same sequence of joint displacements in different contexts. 
For instance, how a grasping action is represented may depend on a relatively distal outcome: on whether (say) it results in the object grasped being eaten or placed. In addition, the joint displacements which realise grasping in one context might in another context realise a different action, such as scratching; or they might fail to realise any action at all (because there is no target object, say). Third, a motor representation of an action outcome locates the action within a hierarchical structure. To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.%
\footnote{On the notion of action outcome in motor representation, see \citet{gallese:2000_inner}, \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}

Why accept that there are motor representations of action outcomes? The first step is to consider evidence that motor processes carry information about action outcomes. For any given marker of motor processing (such as a pattern of neuronal discharge or motor-evoked potentials), how can we test whether that marker carries information about action outcomes? The basic principle is straightforward: vary kinematic and dynamic features while holding constant an action outcome; and, conversely, vary action outcomes while holding kinematic and dynamic features constant. In practice researchers have devised many ingenious ways to achieve this. In order to vary kinematic and dynamic features while holding action outcomes constant, in some studies a single action outcome is achieved using different effectors, hand, mouth or foot, say \citep{rizzolatti:1988_functional,Rizzolatti:2001ug,cattaneo:2010_state-dependent}. A variation on this approach is to contrast performing a grasping action with different tools, so that the same action outcome might require closing or opening the hand depending on the tool used \citep{umilta:2008pliers,cattaneo:2009_representation,rochat:2010_responses}. In order to vary action outcome while holding kinematic and dynamic features constant, researchers have contrasted grasping movements with different distal outcomes such as eating and placing \citep{Fogassi:2005nf,bonini:2010_ventral,cattaneo:2007_impairment}. Another approach is to contrast the same grasping movements performed in the presence or manifest absence of a target object \citep{Umilta:2001zr,villiger:2010_activity}. A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}. 
In each of these cases there is evidence that some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.%
\footnote{ Of course some researchers have raised doubts concerning some of the evidence for the claim that there are markers of motor processing which carry information about action outcomes (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).  On balance, however, the evidence supports this conclusion.}

@@

So far we have argued that motor processes involve representations of action outcomes. It remains to show that such representations ground purposive actions. But this is hardly a further step. How do intentions ground the purposiveness of actions?  On any standard view, an intention represents an outcome, causes an action, and does so in a way that would normally facilitate the outcome's occurrence. Similarly, motor representations of outcomes represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented. To say that motor representations do all this is one way of making precise the metaphor involved in saying that purposive actions are directed to outcomes.  
Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness (as \citealp{pacherie:2008_action}, pp.\ 189-90 has also argued).  
So given the functional roles and contents of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.

@@

We have been comparing actually seeing with imagining seeing and actually acting with imagining acting as a first step towards arguing that visual and motor representations differ in format (which in turn will be background for an argument that motor representation and intention differ in format).  This claim could not easily be established by comparing actually seeing with actually acting because performance differences between these might be explained by bodily or environmental factors only distantly related to the representations involved. By contrast, comparing performance in imagining seeing and imagining acting does provide reasons to conclude that visual and motor representations differ in format. To see why, contrast imagining rotating a ball with imagining seeing a ball rotating.%
\footnote{Imagining acting without also imagining seeing may be difficult in practice, and conversely; it may also sometimes be difficult to distinguish imagining acting from imagining seeing \citep[as][p.\ 170 suggest]{currie:1997_mental}. However ordinary subjects can separate the two well enough to confirm predictions about their differences \citep[see, e.g.,][]{kosslyn:2001_imagining}.
}
As already mentioned, how quickly the former can be done is a function of how long it would take the agent to rotate the ball, whereas how quickly the latter can be done depends on how rapidly the ball can rotate and still be perceived as rotating. Similarly, we mentioned that factors making actually acting more effortful also make imagining acting more effortful. For instance, in some cases rotating a ball clockwise is easier than rotating it anti-clockwise, 
%*correction!
and so is imagining rotating a ball.  By contrast, the effort involved in actually seeing or imagining seeing a ball rotate does not similarly differ depending on direction.  These and other performance differences are plausibly a consequence of a difference in format between motor and visual representations.

@@

The idea that some demonstrative concepts refer to actions by deferring to motor representations immediately raises a question. Could a demonstrative concept really defer to a motor representation? It seems clear that we can't select a motor representation to defer to in the same way that we can select a map when we say `Follow this route!'. After all, motor representations are not things we can point to with our hands. Nevertheless, it does seem that motor representations are available in some sense. To start with an analogy, consider pantomiming an action to yourself. You are rehearsing part of an operation which involves precisely grasping a delicate structure with some tweezers. Just as someone might point to a map and say `Follow that route!', so also they could point to your pantomime and say `Do that!'. In our analogy, the pantomime stands in for a motor representation of the action.  The demonstrative in `Do that!' refers to an action by deferring to the pantomime. Of course this analogy doesn't show that demonstrative concepts can defer to motor representations (at least not unless pantomimes are motor representations of actions). But now consider purely mental pantomime---that is, phenomenologically action-like imagination. 
One might use this kind of imagination to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!'. 
It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in imagining acting.

But is this really possible? Someone might object that in appealing to imagining acting we are sneaking intentions in through the back door. How can we be sure that it is ever really a motor representation rather than an intention that one defers to in thinking `Do that!'?  Contrast two cases of phenomenologically action-like imagination, both involving a tool. In the first case, imagine grasping an object with the tool; and in the second case, imagine releasing the same object with the tool. The two cases are different but in what does their difference consist? 
Could it be a difference in intention that explains how they differ?  
Since it is possible to grasp without intending to grasp, it is surely also possible to imagine grasping without imagining intending to grasp. 
And note that to imagine grasping without imagining intending to grasp does not necessarily require one to imagine non-intentionally grasping. 
(It is easy to miss this point by confusing it with the issue of whether the imagining must itself be intentional. Whether or not it is possible to imagine grasping or releasing without intending to imagine so acting, one can certainly imagine grasping or releasing without  imagining intending to so act.) So to rule out the possibility that the two imaginings differ with respect to intention, let both be neutral with respect to intention. Is the difference between the two cases then due to differing movements or patterns of muscle contraction? To rule this possibility out, let the movements and muscle contractions involved in both cases be as similar as possible: let the difference between grasping and releasing the object be a matter of how the tool is configured rather than of how your body moves. With the contrast cases elaborated by these two stipulations, it is plausible that the difference between the two imaginings will be due to the different motor representations involved in grasping and releasing. So in these cases, the thought `Do that!' will refer to grasping (or releasing) by deferring to a motor representation of grasping (or of releasing). It follows that demonstrative concepts can refer to actions by deferring to motor representations.

@@

At this point we should acknowledge a complication.  Above we argued that experience is needed for the acquisition of demonstrative concepts of ways of acting but not for their continued use in practical reasoning.  This matters because forming an intention does not always involve newly experiencing a way of acting. However, things are more complicated concerning what to act on. Where an action involves particular things to be acted on, agents will not generally have past experiences of these.  Accordingly, demonstrative reference by deference to motor representation requires that some motor representations of particular targets for action are associated with experiences of those targets.  When we spontaneously intend to act on an object of which we have no previous experience (and perhaps in other cases too), our experience of that object would have to depend in part on motor representations of it.  There is indeed some evidence for the ubiquity of motor aspects of experience (and also evidence that motor representation does not always modulate how objects are experienced, but that is not relevant here).  For example, temporarily changing subjects' motor abilities by artificially extending their reach systematically affects their judgements of how far away objects are \citep{linkenauger:2009_effects, costantini2011tool}. Given that such judgements are based on experience,\footnote{
It may be objected that these judgements could reflect non-experiential expectations; \citet[pp.\ 203-4]{witt:2011_action} reviews evidence against this possibility.
} 
this illustrates one way in which experience may be shaped in part by motor representation of objects.  
There is also evidence that experiential judgements of size can be influenced by ability \citep[e.g.][]{witt2009kicking}.  
It is plausible, then, that some motor representations of objects  modulate experiences of them.  
This indicates that it is possible to refer not only to ways of acting by also to what to act on by deferring to motor representations.
